---
title: "36-315 Project [Change Title]"
author: "Alice Le, Michelle Jiang, Hailey Tan, Iris Yin"
output:
  pdf_document:
  html_document:
    toc: yes
    toc_float: yes
    code_folding: show
urlcolor: blue
---

## Motivation section 

What is the quality of these elements? Does the title tell us what to expect of the report to follow? 
Does the motivation section convince the reader that the broader study
of this data set could elicit interesting insights?

## Description of Dataset 

```{r, warning = True, message = F}
library(tidyverse)
library(ggplot2)

shopping_data <- read.csv("shopping_behavior_updated.csv")
head(shopping_data)
```

What is the quality of this description? Does it clearly communicate what
the rows and columns (i.e., observations and variables) are in the dataset in a way that is understandable
to a CMU statistics undergraduate audience?

## 3 Research Questions

What is the quality of the research questions? Are they well-motivated by
real-world/scientific interests? Or are they shallow? For example, a very shallow question would be
in the form, “what does the distribution of this variable look like?” - three questions like this would
guarantee 0pts for this item. A more interesting question would motivate why we would like to inspect
particular distributions with real-world context

1. What factors predict how much a customer spends (Purchase Amount)?

Variables involved:

Age

Gender

Category / Item Purchased

Size, Color

Season

Discount Applied

Promo Code Used

Previous Purchases

Subscription Status

Location

Payment Method

Frequency of Purchases

Possible analyses:

Multiple regression to identify significant predictors of spending

ANOVA to compare spending across categories or seasons

Interaction effects (e.g., Does discount impact high-frequency shoppers more?)

✅ Boxplots

Purchase Amount vs. Gender

These show differences in spending across groups.

✅ Spatial Graph

Average Purchase Amount by Location

✅ Heatmap

Correlations between:

Purchase Amount

Age

Frequency

Previous Purchases

Useful to identify strong numerical predictors.

2. How does season affect the types of items purchased and the color of the items?

Variables involved:

Possible analyses:

Chi-square test for association

Logistic regression (Predict “Promo Code Used = Yes/No”)

Cluster analysis to uncover shopper groups (Loyal + Discount-heavy vs. Casual + Full-price shoppers)

✅ Faceted Bar Chart

Bar Chart of Discount/Promo Code Usage, Faceted by Subscription Status, with Frequency Groups on the X-axis

This shows all 3 variables clearly:

Facet panels = Subscription Status (Subscribed vs Not Subscribed)

X-axis = Frequency of Purchases (Low / Medium / High)

Bar fill = Discount or Promo Code Used (Yes/No)


✅  Scatter plot

Frequency of Purchases (colored by subscription status) vs. purchase amount
If frequency is truly numeric and continuous, this can show trends.


3. How do discounts and shipping speed affect customer satisfaction (Review Rating)?

Variables involved:

Review Rating

Category

Size

Color

Season

Shipping Type

Purchase Amount (could influence expectations)

Possible analyses:

ANOVA (Do different product categories receive different ratings?)

Regression predicting rating

Comparing ratings across seasons or color/size groups

✅ Boxplots

Review Rating vs. Category

Review Rating vs. Season

This shows which categories or seasons perform best.

✅ Bar Chart

Average Review Rating by Color

Average Rating by Size

✅ Scatterplot

Purchase Amount vs. Review Rating
Higher spending may correlate with more satisfaction—or dissatisfaction.

✅ Violin Plot (optional)

For Rating distributions by Category or Season—gives more detail than a boxplot.

\newpage 

## Question 1 

Graph, Description

\newpage 

## Question 2

Graph, Description

\newpage 

## Question 3 

Graph, Description 

\newpage 

## Evaluation/Future work 

***
***


# Old Draft 

```{r, warning = F, message = F}
library(tidyverse)
library(ggplot2)

url <- "https://raw.githubusercontent.com/AliceeLe/36315/main/diabetes_012_health_indicators_BRFSS2015.csv"
diabetes_dataset <- read.csv(url, header = TRUE, stringsAsFactors = FALSE)

head(diabetes_dataset)
```

## EDA

### Alice's EDA 

```{r}
ggplot(diabetes_dataset, aes(x = factor(Diabetes_012), y = BMI, fill = factor(Diabetes_012))) +
  geom_boxplot() +
  facet_wrap(~ HighBP,
             labeller = labeller(HighBP = c("0" = "No High Blood Pressure", "1" = "High Blood Pressure"))) +
  scale_fill_manual(values = c("0" = "lightblue", "1" = "gold", "2" = "salmon"),
                    labels = c("0" = "None", "1" = "Prediabetes", "2" = "Diabetes")) +
  labs(x = "Diagnosis",
       y = "BMI",
       fill = "Condition",
       title = "BMI Distribution by Diagnosis and High Blood Pressure")

```

### Hailey's EDA 
```{r}

diabetes_dataset %>%
  mutate(
    Diabetes_Status = factor(Diabetes_012,
                             levels = c(0, 1, 2),
                             labels = c("No Diabetes", "Prediabetes", "Diabetes")),
    High_BP = factor(HighBP,
                     levels = c(0, 1),
                     labels = c("Normal BP", "High BP"))
  ) %>%
  ggplot(aes(x = BMI, y = GenHlth, color = Diabetes_Status)) +
  geom_point(alpha = 0.4, size = 1.5) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 1.2) +
  scale_color_manual(values = c("No Diabetes" = "green", 
                                 "Prediabetes" = "orange", 
                                 "Diabetes" = "red")) +
  labs(
    title = "Relationship Between BMI and General Health by Diabetes Status",
    subtitle = "Higher general health scores indicate worse health (1=Excellent, 5=Poor)",
    x = "Body Mass Index (BMI)",
    y = "General Health Rating (1-5 scale)",
    color = "Diabetes Status"
  )
```



### Michelle's EDA
```{r}
diabetesByAge <- diabetes_dataset %>%
  mutate(AgeCategory = factor(Age, levels = sort(unique(Age)),
                              labels = c("18–24","25–29","30–34","35–39","40–44","45–49",
                                         "50–54","55–59","60–64","65–69","70–74","75–79","80+")))
ggplot(diabetesByAge, aes(x = BMI,
                 y = AgeCategory,
                 color = factor(HighBP))) +
  geom_point(alpha = 0.4) +
  scale_color_manual(values = c("blue", "red"),
                     labels = c("No High Blood Pressure", "High Blood Pressure")) +
  labs(
    title = "BMI and Age vs. High Blood Pressure Status",
    x = "Body Mass Index (BMI)",
    y = "Age Category",
    color = "High Blood Pressure",
    caption = "Data: Diabetes Health Indicators Dataset"
  )
```

### Iris's EDA 

```{r}

heat_data <- diabetes_dataset %>%
  group_by(Age, Diabetes_012) %>%
  summarize(mean_BMI = mean(BMI))

ggplot(heat_data,
       aes(x = Age, y = Diabetes_012, fill = mean_BMI)) +
  geom_tile() +
  labs(
    title = "Heatmap of Average BMI by Age and Diabetes Status",
    x = "Age (BRFSS numeric categories)",
    y = "Diabetes_012 (0 = None, 1 = Prediabetes, 2 = Diabetes)",
    fill = "Mean BMI"
  )
```