---
title: "36-315 Project [Change Title]"
author: "Alice Le, Michelle Jiang, Hailey Tan, Iris Yin"
output:
  pdf_document:
  html_document:
    toc: yes
    toc_float: yes
    code_folding: show
urlcolor: blue
---

## Motivation section 

What is the quality of these elements? Does the title tell us what to expect of the report to follow? 
Does the motivation section convince the reader that the broader study
of this data set could elicit interesting insights?

## Description of Dataset 

```{r, warning = True, message = F}
library(tidyverse)
library(ggplot2)

shopping_data <- read.csv("shopping_behavior_updated.csv")
head(shopping_data)
```

What is the quality of this description? Does it clearly communicate what
the rows and columns (i.e., observations and variables) are in the dataset in a way that is understandable
to a CMU statistics undergraduate audience?

## 3 Research Questions

What is the quality of the research questions? Are they well-motivated by
real-world/scientific interests? Or are they shallow? For example, a very shallow question would be
in the form, “what does the distribution of this variable look like?” - three questions like this would
guarantee 0pts for this item. A more interesting question would motivate why we would like to inspect
particular distributions with real-world context

1. What factors predict how much a customer spends (Purchase Amount)?

Variables involved:

Age

Gender

Category / Item Purchased

Size, Color

Season

Discount Applied

Promo Code Used

Previous Purchases

Subscription Status

Location

Payment Method

Frequency of Purchases

Possible analyses:

Multiple regression to identify significant predictors of spending

ANOVA to compare spending across categories or seasons

Interaction effects (e.g., Does discount impact high-frequency shoppers more?)

✅ Boxplots - Hailey

Purchase Amount vs. Gender

```{r}
ggplot(shopping_data, aes(x = Gender, y = Purchase.Amount..USD., fill = Gender)) +
  geom_boxplot(alpha = 0.7, outlier.shape = 16, outlier.size = 2) +
  scale_fill_manual(values = c("Male" = "lightblue", "Female" = "lightpink")) +
  labs(
    title = "Purchase Amount Distribution by Gender",
    subtitle = "Comparison of Shopping Spending Patterns",
    x = "Gender",
    y = "Purchase Amount (USD)",
    caption = "Data source: Shopping Behavior Dataset"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 12),
    axis.title = element_text(size = 12),
  ) +
  stat_summary(fun = mean, geom = "point", shape = 23, size = 3, 
               fill = "white", color = "black")
```



✅ Spatial Graph - Hailey

Average Purchase Amount by Location

```{r}
# average purchase amount by location (state)
location_purchase <- shopping_data %>%
  mutate(location_lower = tolower(Location)) %>%
  group_by(location_lower) %>%
  summarize(avg_purchase = mean(Purchase.Amount..USD., na.rm = TRUE))

# state map data
states_map <- map_data("state")

# join the shopping data with map data
map_with_shopping <- states_map %>%
  left_join(location_purchase, by = c("region" = "location_lower"))

# choropleth map
ggplot(map_with_shopping, aes(x = long, y = lat, group = group, fill = avg_purchase)) +
  geom_polygon(color = "white", size = 0.2) +
  coord_fixed(1.3) +
  scale_fill_gradient2(
    low = "#2E7D32",      # Dark green for low purchases
    mid = "#FFF59D",      # Light yellow for medium purchases
    high = "#D32F2F",     # Dark red for high purchases
    midpoint = median(location_purchase$avg_purchase, na.rm = TRUE),
    name = "Avg Purchase\nAmount (USD)",
    na.value = "grey50"
  ) +
  labs(
    title = "Average Purchase Amount by State",
    subtitle = "Shopping Behavior Analysis",
    caption = "Data source: Shopping Behavior Dataset"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 12),
    legend.position = "right"
  )
```


✅ Boxplot - Iris

Correlations between:

Purchase Amount
PUrchase method

Scatterplot - Iris
Purchase Amount vs Age

2. How does season affect the types of items purchased and the color of the items?

Variables involved:

Possible analyses:

Chi-square test for association

Logistic regression (Predict “Promo Code Used = Yes/No”)

Cluster analysis to uncover shopper groups (Loyal + Discount-heavy vs. Casual + Full-price shoppers)

✅ Word Cloud - Michelle

color frequency by season


✅ Heatmap - Michelle

This handles large numbers of categories beautifully.

Why it works:

Items go on the Y-axis

Seasons go on the X-axis

Color = number of purchases

It scales well even if you have 20–50+ item types.

Interpretation:

Darker = purchased more often

Lighter = purchased less often
Frequency of Purchases (colored by subscription status) vs. purchase amount
If frequency is truly numeric and continuous, this can show trends.

✅ 1. Chi-Square Test of Independence (Most appropriate)

Use this when both variables are categorical.

You can run:
a. Season × Item Purchased

Tests whether certain items are bought more often in some seasons.

b. Season × Color

Tests whether certain colors are bought more often in certain seasons.


3. How do discounts and shipping speed affect customer satisfaction (Review Rating)?

Variables involved:

Review Rating

Category

Size

Color

Season

Shipping Type

Purchase Amount (could influence expectations)

Possible analyses:

ANOVA (Do different product categories receive different ratings?)

Regression predicting rating

Comparing ratings across seasons or color/size groups

⭐ 2. Violin Plot - Alice

A more detailed version of the boxplot.

Why use it:

Shows full distribution of review ratings

Useful if ratings aren’t normally distributed

Same structure as boxplot:

X = Shipping Type

Y = Review Rating

Color = Discount


- A dendrogram is an excellent choice for your question, because it shows clusters of customers based on: - Alice

Discount used (Yes/No)

Shipping type (categorical)

Review rating (numeric or binned)
\newpage 

## Question 1 

Graph, Description

\newpage 

## Question 2

Graph, Description

\newpage 

## Question 3 

Graph, Description 

\newpage 

## Evaluation/Future work 

***
***


# Old Draft 

```{r, warning = F, message = F}
library(tidyverse)
library(ggplot2)

url <- "https://raw.githubusercontent.com/AliceeLe/36315/main/diabetes_012_health_indicators_BRFSS2015.csv"
diabetes_dataset <- read.csv(url, header = TRUE, stringsAsFactors = FALSE)

head(diabetes_dataset)
```

## EDA

### Alice's EDA 

```{r}
ggplot(diabetes_dataset, aes(x = factor(Diabetes_012), y = BMI, fill = factor(Diabetes_012))) +
  geom_boxplot() +
  facet_wrap(~ HighBP,
             labeller = labeller(HighBP = c("0" = "No High Blood Pressure", "1" = "High Blood Pressure"))) +
  scale_fill_manual(values = c("0" = "lightblue", "1" = "gold", "2" = "salmon"),
                    labels = c("0" = "None", "1" = "Prediabetes", "2" = "Diabetes")) +
  labs(x = "Diagnosis",
       y = "BMI",
       fill = "Condition",
       title = "BMI Distribution by Diagnosis and High Blood Pressure")

```

### Hailey's EDA 
```{r}

diabetes_dataset %>%
  mutate(
    Diabetes_Status = factor(Diabetes_012,
                             levels = c(0, 1, 2),
                             labels = c("No Diabetes", "Prediabetes", "Diabetes")),
    High_BP = factor(HighBP,
                     levels = c(0, 1),
                     labels = c("Normal BP", "High BP"))
  ) %>%
  ggplot(aes(x = BMI, y = GenHlth, color = Diabetes_Status)) +
  geom_point(alpha = 0.4, size = 1.5) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 1.2) +
  scale_color_manual(values = c("No Diabetes" = "green", 
                                 "Prediabetes" = "orange", 
                                 "Diabetes" = "red")) +
  labs(
    title = "Relationship Between BMI and General Health by Diabetes Status",
    subtitle = "Higher general health scores indicate worse health (1=Excellent, 5=Poor)",
    x = "Body Mass Index (BMI)",
    y = "General Health Rating (1-5 scale)",
    color = "Diabetes Status"
  )
```



### Michelle's EDA
```{r}
diabetesByAge <- diabetes_dataset %>%
  mutate(AgeCategory = factor(Age, levels = sort(unique(Age)),
                              labels = c("18–24","25–29","30–34","35–39","40–44","45–49",
                                         "50–54","55–59","60–64","65–69","70–74","75–79","80+")))
ggplot(diabetesByAge, aes(x = BMI,
                 y = AgeCategory,
                 color = factor(HighBP))) +
  geom_point(alpha = 0.4) +
  scale_color_manual(values = c("blue", "red"),
                     labels = c("No High Blood Pressure", "High Blood Pressure")) +
  labs(
    title = "BMI and Age vs. High Blood Pressure Status",
    x = "Body Mass Index (BMI)",
    y = "Age Category",
    color = "High Blood Pressure",
    caption = "Data: Diabetes Health Indicators Dataset"
  )
```

### Iris's EDA 

```{r}

heat_data <- diabetes_dataset %>%
  group_by(Age, Diabetes_012) %>%
  summarize(mean_BMI = mean(BMI))

ggplot(heat_data,
       aes(x = Age, y = Diabetes_012, fill = mean_BMI)) +
  geom_tile() +
  labs(
    title = "Heatmap of Average BMI by Age and Diabetes Status",
    x = "Age (BRFSS numeric categories)",
    y = "Diabetes_012 (0 = None, 1 = Prediabetes, 2 = Diabetes)",
    fill = "Mean BMI"
  )
```