---
title: "Understanding Retail Consumer Behavior: Predictors of Spending, Product Choice, and Customer Satisfaction"
author: "Alice Le, Michelle Jiang, Hailey Tan, Iris Yin"
output:
  pdf_document:
  html_document:
    toc: yes
    toc_float: yes
    code_folding: show
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,      # hide all R code
  message = FALSE,   # hide messages
  warning = FALSE,   # hide warnings
  results = "hide"   # hide printed output unless overridden
)
```

# Motivation section (Iris)

What is the quality of these elements? Does the title tell us what to expect of the report to follow? 
Does the motivation section convince the reader that the broader study
of this data set could elicit interesting insights?

# Description of Dataset (Alice)

```{r, message = F}
library(tidyverse)
library(ggplot2)

shopping_data <- read.csv("shopping_behavior_updated.csv")
head(shopping_data)
```


As stated, the dataset describes the shopping behavior of consumers, where each row represents an individual customer and their purchasing activity, including what they bought, how much they spent, how often they shop, their demographic attributes, and whether they used discounts or promotional codes.


Generally, the dataset is of good quality with no missing values or duplicate rows. There are 3900 rows in total, and each row corresponds to a unique customer’s purchase transaction collected across a variety of U.S. states.


**Where does it come from??**
As stated, the dataset describes the shopping behavior of consumers, where [insert]

The variables in the dataset are:

+ `Customer.ID`: The unique ID of each customer
+ `Age`: The age of customer (in years). Range from 18 to 70 years old.
+ `Gender`: Gender of customer
+ `Item.Purchased`: Type of item that the customer bought, such as Blouse, Sweater, Scarf, etc.
+ `Category`: Broader category of the item purchased. There are 4 categories: Clothing, Footwear, Outerwear, and Accessories
+ `Purchase.Amount..USD`: The amount spent on the purchased item, in US dollars
+ `Location`: The US state where the customer resides
+ `Size`: Size of the clothing item. There are 4 sizes: S, M, L, XL
+ `Color`: Color of the clothing item
+ `Season`: 
+ `Review.Rating`
+ `Subscription.Status`
+ `Shipping.Type`: The method of shipping used for the order (e.g., "Express", "Free Shipping", "Next Day Air", "2-Day Shipping", "Store Pickup"). Indicates speed and delivery preferences
+ `Discount.Applied`: Whether a discount was applied to the purchase
+ `Promo.Code.Used`: Whether a promotion code was applied to the purchase
+ `Previous.Purchases`: The number of purchases the customer made before this transaction
+ `Payment.Method`: The payment method used for the transaction. There are 6 options: Cash, Bank Transfer, Debit Card, Credit Card, Venmo, and Paypal 
+ `Frequency.of.Purchases`: How often the customer shops. The frequencies include Weekly, Fortnightly, Bi-weekly, Monthly, Every 3 Months, Quarterly, and Annually

# 3 Research Questions 

Stated below are our 3 research questions: 

1. What factors predict how much a customer spends (Purchase Amount)?

2. How does season affect the types of items purchased and the color of the items? 

3. How do discounts and shipping speed affect customer satisfaction (Review Rating)?

These research questions address core mechanisms of consumer behavior: spending determinants, seasonal purchasing patterns, and satisfaction drivers. Together, they allow us to examine how economic, temporal, and operational factors shape shopping outcomes in a way that is meaningful for both behavioral analysis and retail decision-making.


\newpage
# Question 1: What factors predict how much a customer spends (Purchase Amount)?

## Boxplots - Hailey

Purchase Amount vs. Gender

```{r}
ggplot(shopping_data, aes(x = Gender, y = Purchase.Amount..USD., fill = Gender)) +
  geom_boxplot(alpha = 0.7, outlier.shape = 16, outlier.size = 2) +
  scale_fill_manual(values = c("Male" = "lightblue", "Female" = "lightpink")) +
  labs(
    title = "Purchase Amount Distribution by Gender",
    subtitle = "Comparison of Shopping Spending Patterns",
    x = "Gender",
    y = "Purchase Amount (USD)",
    caption = "Data source: Shopping Behavior Dataset"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 12),
    axis.title = element_text(size = 12),
  ) +
  stat_summary(fun = mean, geom = "point", shape = 23, size = 3, 
               fill = "white", color = "black")
```



## Spatial Graph - Hailey

Average Purchase Amount by Location

```{r}
# average purchase amount by location (state)
location_purchase <- shopping_data %>%
  mutate(location_lower = tolower(Location)) %>%
  group_by(location_lower) %>%
  summarize(avg_purchase = mean(Purchase.Amount..USD., na.rm = TRUE))

# state map data
states_map <- map_data("state")

# join the shopping data with map data
map_with_shopping <- states_map %>%
  left_join(location_purchase, by = c("region" = "location_lower"))

# choropleth map
ggplot(map_with_shopping, aes(x = long, y = lat, group = group, fill = avg_purchase)) +
  geom_polygon(color = "white", size = 0.2) +
  coord_fixed(1.3) +
  scale_fill_gradient2(
    low = "#2E7D32",      # Dark green for low purchases
    mid = "#FFF59D",      # Light yellow for medium purchases
    high = "#D32F2F",     # Dark red for high purchases
    midpoint = median(location_purchase$avg_purchase, na.rm = TRUE),
    name = "Avg Purchase\nAmount (USD)",
    na.value = "grey50"
  ) +
  labs(
    title = "Average Purchase Amount by State",
    subtitle = "Shopping Behavior Analysis",
    caption = "Data source: Shopping Behavior Dataset"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 12),
    legend.position = "right"
  )
```


## Boxplot - Iris

Correlations between:

Purchase Amount
PUrchase method

```{r}
library(ggplot2)
ggplot(shopping_data, aes(x = Payment.Method, y = Purchase.Amount..USD., fill= Payment.Method)) +
  geom_boxplot() +
  labs(
    title = "Purchase Amount by Purchase Method",
    x = "Purchase Method",
    y = "Purchase Amount (USD)"
  )

```

## Scatterplot - Iris
Purchase Amount vs Age

```{r}
avg_by_age <- shopping_data %>%
  group_by(Age) %>%
  summarise(
    mean_purchase = mean(Purchase.Amount..USD., na.rm = TRUE),
    n = n()
  )

ggplot(avg_by_age, aes(x = Age, y = mean_purchase, color = mean_purchase)) +
  geom_point(size = 2) +
  scale_color_gradient(
    low  = "blue",   # cheaper
    high = "red"     # higher
  ) +
  labs(
    title = "Average Purchase Amount by Age",
    x = "Age",
    y = "Mean Purchase Amount (USD)",
    color = "Mean\nPurchase"
  )

```

\newpage 

## Question 2: How does season affect the types of items purchased and the color of the items? 

**Season does not seem to have a significant affect on the types of items purchased and color of the items.**

**Variables involved: Season (Winter, Spring, Summer, Fall).Type of item purchased – e.g., Category or Item Purchased, Color of item – Color**

**Possible analyses:Use visualizations such as stacked/clustered bar charts or mosaic plots to compare item types and colors across seasons.**

Chi-square test for association

Logistic regression (Predict “Promo Code Used = Yes/No”)


```{r}

df <- shopping_data %>%
  mutate(
    promo_used = if_else(Promo.Code.Used == "Yes", 1, 0)
  )

# Fit logistic regression
promo_model <- glm(
  promo_used ~ Age +
    Purchase.Amount..USD. +
    Season +
    Discount.Applied +
    Payment.Method +
    Subscription.Status,
  data = df,
  family = binomial
)

summary(promo_model)

```



## Word Cloud - Michelle

color frequency by season

```{r}
# PUT YOUR CODE AND GRAPH HERE
library(tidyverse)
library(tm)
library(wordcloud)

shopping <- read.csv("shopping_behavior_updated.csv")
season_text <- shopping %>%
  group_by(Season) %>%
  summarise(text = paste(Color, collapse = " "))

corpus <- VCorpus(VectorSource(season_text$text))

corpus <- corpus %>%
  tm_map(content_transformer(tolower)) %>%
  tm_map(removePunctuation) %>%
  tm_map(removeNumbers) %>%
  tm_map(removeWords, stopwords("english"))

tdm <- TermDocumentMatrix(corpus)

tdm_matrix <- as.matrix(tdm)
colnames(tdm_matrix) <- season_text$Season

comparison.cloud(tdm_matrix,random.order = FALSE, colors = c("goldenrod", "palegreen2", "pink", "skyblue4"), title.size = 2)


```


## Heatmap - Michelle

This handles large numbers of categories beautifully.

Why it works:

Items go on the Y-axis

Seasons go on the X-axis

Color = number of purchases

It scales well even if you have 20–50+ item types.

Interpretation:

Darker = purchased more often

Lighter = purchased less often
Frequency of Purchases (colored by subscription status) vs. purchase amount
If frequency is truly numeric and continuous, this can show trends.

```{r}
shopping_data_counts <- shopping_data %>% 
  mutate(Season = factor(Season, 
                         levels = c("Spring", "Summer", "Fall", "Winter"))) %>%
  count(Item.Purchased, Season, name = "purchase_count")


ggplot(shopping_data_counts, aes(x = Season, y = Item.Purchased, fill = purchase_count)) +
  geom_tile(color = "white") +
  scale_fill_gradient(low = "lightyellow", high = "darkred") +
  labs(
    title = "Purchases by Item and Season",
    x = "Season",
    y = "Item",
    fill = "Number of Purchases"
  )

```

### 1. Chi-Square Test of Independence (Most appropriate)

Use this when both variables are categorical.

You can run:
a. Season × Item Purchased

Tests whether certain items are bought more often in some seasons.

b. Season × Color

Tests whether certain colors are bought more often in certain seasons.

\newpage 

# Question 3: How do discounts and shipping speed affect customer satisfaction (Review Rating)?

Description: 

## Violin plot  - Alice

```{r}
ggplot(shopping_data, aes(x = Shipping.Type, y = Review.Rating, fill = Shipping.Type)) +
  geom_violin(trim = FALSE, alpha = 0.7, color = NA) +
  geom_boxplot(width = 0.18, outlier.size = 0.8, alpha = 0.9) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title = "Review Ratings by Shipping Type",
    x = "Shipping Type",
    y = "Review Rating"
  ) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 30, hjust = 1)
  )
```

```{r, echo=FALSE, results='show'}
anova_model <- aov(Review.Rating ~ Shipping.Type, data = shopping_data)
summary(anova_model)
```

```{r}
ggplot(shopping_data, aes(x = Discount.Applied, y = Review.Rating, fill = Discount.Applied)) +
  geom_violin(trim = FALSE, alpha = 0.7, color = NA) +
  geom_boxplot(width = 0.18, outlier.size = 0.8, alpha = 0.9) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title = "Review Ratings by Discount Usage",
    x = "Discount Applied?",
    y = "Review Rating"
  ) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 30, hjust = 1)
  )

```

```{r, echo=FALSE, results='show'}
t.test(Review.Rating ~ Discount.Applied, data = shopping_data)
```

\newpage 


# Evaluation/Future work (Michelle)



